geochemistry

### 2025-01-29 geochemistry.qmd
Average table and graphs for relevant elements

library(readr)
library(dplyr)

library(janitor)
library(tidyr)

@@

pxrf_dataset <- pxrf_raw |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type", "Application", "Method",
         "ElapsedTime",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2",
         "V", "Cr", "MnO", "Fe2O3", "Co", "Ni", "Cu", "Zn", "As", "Se", "Rb",
         "Sr", "Y", "Zr", "Nb", "Mo", "Rh", "Ag", "Cd", "Sn", "Sb", "Ba", "La",
         "Ce", "Hf", "Ta", "W", "Pt", "Au", "Hg", "Tl", "Pb", "Bi", "Th", "U",
         "Cal Check") |> 
  mutate("MgO_detect" = case_when(
    MgO == 0 | MgO == "< LOD" ~ "LOD",
    TRUE ~ "Value"))

pxrf_analysis <- pxrf_dataset |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb") # Removed elements with only 0 or <LOD values in all samples

@@

```{r}
#| label: tbl-occurrences
#| tbl-cap: A table.

# Example dataset
data <- data.frame(
  Group = c("A", "A", "B", "B", "A", "B", "C", "C", "C", "A"),
  Category1 = c("X", "Y", "X", "Y", "X", "Y", "X", "X", "Y", "Y"),
  Category2 = c("P", "P", "Q", "Q", "P", "P", "Q", "P", "Q", "P")
)

# Combine the multiple columns into a single table
combined_table <- pxrf_analysis |> 
  mutate(across(
    .cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb"),
    .fns = ~ case_when(
      . == 0 | . == "< LOD" ~ "No",
      . > 0 ~ "Yes",
      TRUE ~ as.character(.))
    )) |> 
  mutate(`Chert type` = case_when(
    `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "TYPE 2",
    `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "TYPE 6",
    TRUE ~ `Formation/Type`))

@@

  select("Chert type", Yes) |> 
  rename(Au = Yes)
# Pb
tbl_Pb <- combined_table |> 
  tabyl(`Chert type`, Pb) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Pb = Yes)

@@

library(gt)

combined_table |> 
  select(`Formation/Type`, MgO, Al2O3) |> 
  t

# Print the resulting table
print(combined_table)

>>>>

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)

@@

pxrf_analysis <- pxrf_raw |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr") |>  # Removed elements with 0 or <LOD values in all samples or in high percentages in all groups
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2",
    `Formation/Type` == "TYPE 3" ~ "T3",
    `Formation/Type` == "TYPE 4" ~ "T4",
    `Formation/Type` == "TYPE 5" ~ "T5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  ))

@@

```{r}
#| label: tbl-occurrences
#| tbl-cap: A table.
#| include: false

# Table to see the occurrence of 0 or <LOD values within each element and by rm type, to choose the elements in which to apply statistics.
# Combine the multiple columns into a single table
combined_table <- pxrf_raw |>
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb") |> 
  mutate(across(
    .cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb"),
    .fns = ~ case_when(
      . == 0 | . == "< LOD" ~ "No",
      . > 0 ~ "Yes",
      TRUE ~ as.character(.))
    )) |> 
  mutate(`Chert type` = case_when(
    `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "TYPE 2",
    `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "TYPE 6",
    TRUE ~ `Formation/Type`))

@@

  select("Chert type", Yes) |> 
  rename(Au = Yes)
# Pb
tbl_Pb <- combined_table |> 
  tabyl(`Chert type`, Pb) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>
  select("Chert type", Yes) |> 
  rename(Pb = Yes)

@@

combined_tbl[, 2:26] <- lapply(combined_tbl[, 2:26], function(x) as.numeric(sub("%", "", x)))
gt(combined_tbl) %>%
  data_color(
    columns = vars("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb"),  # Multiple columns
    colors = scales::col_numeric(
      palette = c("#f8f0fc", "#dab6fc", "#b6f7c1"), # Gradient from red to green
      domain = c(0, 100))
  )

```{r}
#| label: averages

pxrf_average <- pxrf_analysis |> 
  select(-Sample_ID, -pXRF_ID) |> 
  mutate(across(-`Formation/Type`, ~ as.numeric(.))) |>    # Convert valid numbers
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .))) |>  # Replace NAs with 0
  group_by(`Formation/Type`) |> 
  summarise(across(where(is.numeric), ~ round(mean(.), 3)))
  pivot_wider(names_from = `Formation/Type`,
              values_from = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr"))

average_MgO <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = MgO)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "MgO")

average_Al2O3 <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = Al2O3)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "Al2O3")

average_SiO2 <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = SiO2)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "SiO2")

average_P2O5 <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = P2O5)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "P2O5")

average_S <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = S)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "S")

average_Cl <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = Cl)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "Cl")

average_K2O <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = K2O)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "K2O")

average_CaO <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = CaO)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "CaO")

average_TiO2 <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = TiO2)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "TiO2")

average_Fe2O3 <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = Fe2O3)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "Fe2O3")

average_Sr <- pxrf_average |> 
  ggplot(aes(x = `Formation/Type`, y = Sr)) +
  geom_bar(stat = "identity") +
  theme_minimal() + 
  labs(x = " ",y = "Sr")
